---
title: "Cohort Definitions in OHDSI Phenotype Library"
author: "Gowtham A. Rao"
date: "`r Sys.Date()`"
output:
  html_document:
    number_sections: yes
    toc: no  
  pdf_document: 
    toc: no     
vignette: >
  %\VignetteIndexEntry{Cohort Definitions in OHDSI Phenotype Library}
  %\VignetteEncoding{UTF-8}    
  %\VignetteEngine{knitr::rmarkdown}
---

```{r, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(
  cache = FALSE,
  comment = "#>",
  error = FALSE,
  tidy = FALSE)
```

[PhenotypeLibrary](https://github.com/OHDSI/PhenotypeLibrary) is part of [HADES](https://ohdsi.github.io/Hades/)

# Introduction:

This document shows a summary of changes to the Cohort Definitions in [OHDSI Phenotype Library](https://github.com/OHDSI/PhenotypeLibrary) (Library) over various versions.


* **Cohort Id:** Id used by the [Library](https://github.com/OHDSI/PhenotypeLibrary) to identify the cohort. This is guaranteed to be backward compatible and so may be referenced in OHDSI HADES packages and studies using a combination of version number and cohort id.

* **Cohort Name:** The name of the Cohort Definition

* **Get Results:** Is this part of the [ohdsi-studies/PhenotypeLibraryDiagnostics] (https://github.com/ohdsi-studies/PhenotypeLibraryDiagnostics) study package. If yes, then the results maybe found on [data.ohdsi.org/PhenotypeLibrary](https://data.ohdsi.org/PhenotypeLibrary/)

* **Added Version:** The version of the PhenotypeLibrary R package when the cohort definition was added.

* Added Date: The date the cohort was added. This is usually from [atlas-phenotype.ohdsi.org](https://atlas-phenotype.ohdsi.org/#/home) when the source is [atlas-phenotype.ohdsi.org](https://atlas-phenotype.ohdsi.org/#/home)

* **Added Notes:** Any notes entered by the [OHDSI Phenotype Development and Evaluation Work group](https://teams.microsoft.com/l/team/19%3a66a11855657c46918723073bff9e36f1%40thread.tacv2/conversations?groupId=b464dca0-06d1-477c-b66d-11827e4d0412&tenantId=a30f0094-9120-4aab-ba4c-e5509023b2d5) (work group) for this Cohort Definition.
Deprecated Version: If there is an entry in this field it means the Cohort Definition is deprecated. It will have the version of PhenotypeLibrary when the Cohort Definition was deprecated.

* **Deprecated Date:** The date the Cohort was deprecated. This is usually from [atlas-phenotype.ohdsi.org](https://atlas-phenotype.ohdsi.org/#/home) when the source is [atlas-phenotype.ohdsi.org](https://atlas-phenotype.ohdsi.org/#/home)

* **Deprecated Notes:** Any notes from the [Work group](https://teams.microsoft.com/l/team/19%3a66a11855657c46918723073bff9e36f1%40thread.tacv2/conversations?groupId=b464dca0-06d1-477c-b66d-11827e4d0412&tenantId=a30f0094-9120-4aab-ba4c-e5509023b2d5) regarding deprecation. 
Updated Version: This field indicates the cohort definition has been modified. As a principle a Cohort Definition, once published to the OHDSI Phenotype Library is never updated or modified. This should be a rare occurrence, for situations like human accidental errors.

* **Updated Date:** The date the Cohort was updated.

* **Updated Notes:** Any notes from the [Work group](https://teams.microsoft.com/l/team/19%3a66a11855657c46918723073bff9e36f1%40thread.tacv2/conversations?groupId=b464dca0-06d1-477c-b66d-11827e4d0412&tenantId=a30f0094-9120-4aab-ba4c-e5509023b2d5) regarding update.



```{r tidy=FALSE,eval=TRUE}

  camelCaseToTitleCase <- function(string) {
    errorMessages <- checkmate::makeAssertCollection()
    checkmate::assertCharacter(string, add = errorMessages)
    checkmate::reportAssertions(collection = errorMessages)
  
    string <- gsub("([A-Z])", " \\1", string)
    string <- gsub("([a-z])([0-9])", "\\1 \\2", string)
    substr(string, 1, 1) <- toupper(substr(string, 1, 1))
    return(string)
  }

  report <- PhenotypeLibrary::getPhenotypeLog()
  colnames(report) <- camelCaseToTitleCase(colnames(report))
  opts <- options(knitr.kable.NA = "", format = "latex", escape = TRUE)
  knitr::kable(x = report)
```
