---
title: "Cohort Definitions in OHDSI Phenotype Library"
author: "Gowtham A. Rao"
date: "`r Sys.Date()`"
output:
  html_document:
    number_sections: yes
    toc: no     
vignette: >
  %\VignetteIndexEntry{Cohort Definitions in OHDSI Phenotype Library}
  %\VignetteEncoding{UTF-8}    
  %\VignetteEngine{knitr::rmarkdown}
always_allow_html: true
---

```{r, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(
  cache = FALSE,
  comment = "#>",
  error = FALSE,
  tidy = FALSE)

library(reactable)
```

[PhenotypeLibrary](https://github.com/OHDSI/PhenotypeLibrary) is part of [HADES](https://ohdsi.github.io/Hades/). This document is a summary of changes to the Cohort Definitions in [OHDSI Phenotype Library](https://github.com/OHDSI/PhenotypeLibrary) (Library).



```{r tidy=FALSE,eval=TRUE, echo = FALSE}

camelCaseToTitleCase <- function(string) {
  errorMessages <- checkmate::makeAssertCollection()
  checkmate::assertCharacter(string, add = errorMessages)
  checkmate::reportAssertions(collection = errorMessages)
  
  string <- gsub("([A-Z])", " \\1", string)
  string <- gsub("([a-z])([0-9])", "\\1 \\2", string)
  substr(string, 1, 1) <- toupper(substr(string, 1, 1))
  return(string)
}

report <- PhenotypeLibrary::getPhenotypeLog()
colnames(report) <- camelCaseToTitleCase(colnames(report))

columnDefinitions <- list()

colnamesInReport <- colnames(report)

for (i in (1:length(colnamesInReport))) {
  columnDefinitions[[colnamesInReport[[i]]]] <-
    reactable::colDef(
      name = colnamesInReport[[i]],
      minWidth = nchar(colnamesInReport[[i]]) * 10,
      sortable = TRUE,
      resizable = TRUE,
      show = TRUE,
      html = TRUE,
      na = "",
      align = "left"
    )
}

columnDefinitions[['Cohort Name']] <-
  reactable::colDef(
    name = 'Cohort Name',
    minWidth = 500,
    sortable = TRUE,
    resizable = TRUE,
    show = TRUE,
    html = TRUE,
    na = "",
    align = "left"
  )

columnDefinitions[['Cohort Id']] <-
  reactable::colDef(
    name = 'Cohort Id',
    minWidth = 100,
    sticky = "left",
    sortable = TRUE,
    resizable = TRUE,
    show = TRUE,
    html = TRUE,
    na = "",
    align = "left"
  )

  
```

```{r tidy=FALSE,eval=TRUE, echo = FALSE}

reactable::reactable(
    data = report,
    columns = columnDefinitions, 
    sortable = TRUE,
    resizable = TRUE,
    searchable = TRUE,
    pagination = TRUE,
    showPagination = TRUE,
    showPageInfo = TRUE,
    striped = TRUE,
    compact = FALSE,
    wrap = FALSE,
    showSortIcon = TRUE,
    showSortable = TRUE,
    fullWidth = TRUE,
    showPageSizeOptions = TRUE,
    pageSizeOptions = c(10, 20, 50, 100, 1000),
    defaultPageSize = 100
  )
  
```

* **Cohort Id:** Id used by the [Library](https://github.com/OHDSI/PhenotypeLibrary) to identify the cohort. This is guaranteed to be backward compatible and so may be referenced in OHDSI HADES packages and studies using a combination of version number and cohort id.

* **Cohort Name:** The name of the Cohort Definition
  - [D] Cohort has been deprecated, and is replaced by another cohort. This was a peer reviewed cohort.
  - [W] Cohort was withdrawn by the submitter during peer review. This is not a peer reviewed cohort.
  - [E] Cohort is known to have an error. This is usually not a peer reviewed cohort. Occasionally, a peer reviewed cohort may subsequently be found to have an error and be flagged to [E]
  - [P] Cohort is undergoing peer review.

* **Get Results:** Is this part of the [ohdsi-studies/PhenotypeLibraryDiagnostics](https://github.com/ohdsi-studies/PhenotypeLibraryDiagnostics) study package. If yes, then the results maybe found on [data.ohdsi.org/PhenotypeLibrary](https://data.ohdsi.org/PhenotypeLibrary/)

* **Added Version:** The version of the PhenotypeLibrary R package when the cohort definition was added.

* **Added Date:** The date the cohort was added. This is usually from [atlas-phenotype.ohdsi.org](https://atlas-phenotype.ohdsi.org/#/home) when the source is [atlas-phenotype.ohdsi.org](https://atlas-phenotype.ohdsi.org/#/home)

* **Added Notes:** Any notes entered by the [OHDSI Phenotype Development and Evaluation Work group](https://teams.microsoft.com/l/team/19%3a66a11855657c46918723073bff9e36f1%40thread.tacv2/conversations?groupId=b464dca0-06d1-477c-b66d-11827e4d0412&tenantId=a30f0094-9120-4aab-ba4c-e5509023b2d5) (work group) for this Cohort Definition.
Deprecated Version: If there is an entry in this field it means the Cohort Definition is deprecated. It will have the version of PhenotypeLibrary when the Cohort Definition was deprecated.

* **Deprecated Date:** The date the Cohort was deprecated. This is usually from [atlas-phenotype.ohdsi.org](https://atlas-phenotype.ohdsi.org/#/home) when the source is [atlas-phenotype.ohdsi.org](https://atlas-phenotype.ohdsi.org/#/home)

* **Deprecated Notes:** Any notes from the [Work group](https://teams.microsoft.com/l/team/19%3a66a11855657c46918723073bff9e36f1%40thread.tacv2/conversations?groupId=b464dca0-06d1-477c-b66d-11827e4d0412&tenantId=a30f0094-9120-4aab-ba4c-e5509023b2d5) regarding deprecation. 
Updated Version: This field indicates the cohort definition has been modified. As a principle a Cohort Definition, once published to the OHDSI Phenotype Library is never updated or modified. This should be a rare occurrence, for situations like human accidental errors.

* **Updated Date:** The date the Cohort was updated.

* **Updated Notes:** Any notes from the [Work group](https://teams.microsoft.com/l/team/19%3a66a11855657c46918723073bff9e36f1%40thread.tacv2/conversations?groupId=b464dca0-06d1-477c-b66d-11827e4d0412&tenantId=a30f0094-9120-4aab-ba4c-e5509023b2d5) regarding update.



--------------------
* **Grandfathered status:** Certain cohorts were present in the OHDSI Phenotype library prior to version 3. These cohorts are labeled as being grandfathered in. If the cohorts are identified by the prefix [P] that means, they have not undergone peer review. After peer review and acceptance, the prefix [P] will be removed. Reason for grandfathered status: These cohort definitions did not have the submission requirements specified by the OHDSI Phenotype Development and Evaluation work group.


